---
title: "Qualitative movement analysis for the WB"
output: html_document
---

#Questions:

##Space use variability among speices in the WB
###  Species space use
1. Do species use different rivers?
2. Do species use different sections?
+ Capture data
+ wanding data?

###Species movement patterns
Do species move differently?
1. Movements among tribs
2. Movements based on captures
3. Movements based on antennas? (prob not)
4. Do the density/biomasses of the donor and recip sections matter?



```{r libraries}
library(ggplot2)
library(tidyr)
library(dplyr)
library(lubridate)
library(gridExtra)
```

```{r load raw data}
load(file='coreData')

graphsDir <- './graphs/'
```


```{r a look at raw data}


ggplot(cd[ cd$inside ,], aes( section ) ) +
  geom_freqpoly( aes(color=species),binwidth=1.1 ) +
  scale_y_continuous(limits=c( 0.1,1500 )) +
  facet_wrap( ~river )
```

```{r Do species use different rivers?}

#simple count of obs by spp and river
count <- cd %>% group_by(species,river) %>% tally() %>% bind_rows(data.frame(species=c("ats","ats","bnt"),river=c("wb mitchell","wb obear","wb obear"),n=c(0,0,0)))    

# bind_rows undoes the ordered factor - need to redo to get right order for graph
count$riverOrdered <- factor(count$river,labels=c("WB","OL",'OS',"IL"), levels = c("west brook","wb jimmy","wb mitchell","wb obear"), ordered=T)

countByRiverSpecies <- 
  ggplot( count, aes(riverOrdered,n) ) +
    geom_bar(aes(fill=species), position = "dodge",stat="identity") +
    scale_fill_grey(name = "Species") +
      scale_x_discrete('River') +
      scale_y_continuous("Count") +
      theme_bw(base_size=25) +
      theme( legend.position=c(0.9,0.9) )

ggsave( file=paste0(graphsDir,'countByRiverSpecies.png'), plot=countByRiverSpecies, dpi=600 , width=6,height=5, units='in', scale=2 )

```

Calculate movement
```{r calculate and plot movement Tribs}

cd <- cd %>%
        group_by(tag) %>%
        mutate( sectionLagged = lead( section ),
                detectionDateLagged = lead( detectionDate ),
                riverOrderedLagged = lead( riverOrdered ),
                moveDist = (section-sectionLagged) * 20,
                moveTime = as.numeric((difftime(detectionDateLagged,detectionDate,units="days"))) / 30.5 , # in 'months'
                moveRate = moveDist/moveTime, 
                fromTo = paste(riverOrdered,riverOrderedLagged,sep="_")) %>%
   #     filter( moveTime >= 1 & moveTime < 365 ) %>%
        mutate( month = as.numeric( month( detectionDate ) ) )

table(cd$species,cd$fromTo)
```                

Plot movement
```{r plot movement1}

# graph moves based on the proportion of fish from the 'from' river/species
rCount <- 
  filter( cd, !is.na( riverOrderedLagged ) ) %>%
  count(riverOrdered,species) 

rrCount <- 
filter( cd, !is.na( riverOrderedLagged ) ) %>%
  group_by(riverOrdered, riverOrderedLagged, species) %>%
  tally() %>%
  left_join( .,rCount, by = c("riverOrdered","species") ) %>%
  mutate( prop = n.x / n.y ) %>%
  complete( riverOrdered,riverOrderedLagged,species ) %>%  # not sure why this adds 4* the # of rows
  distinct() 
  #mutate( n.y = ifelse( is.na( n.y ), 0, n.y ) )
  
propMove <- 
  ggplot( rrCount, aes( x=riverOrderedLagged, y=prop ) ) +
    geom_bar( stat = "identity" ) +
    scale_fill_grey(name = "Season") +
    scale_x_discrete('Second river') +
    scale_y_continuous("Proportion") +
    theme_bw(base_size=25) +
    theme( legend.position=c(0.95,0.9) ) +
    geom_text( aes(label=(n.y), x=4.2,y=1.025 ), size=6, family="sans" ) +
    facet_grid( species ~ riverOrdered )  

  ggsave( file=paste0(graphsDir,'propMove.png'), plot=propMove, dpi=600 , width=9,height=5, units='in', scale=2 )
  

```


Plot seasonal movement
```{r plot movement}

# graph moves based on the proportion of fish from the 'from' river/species
rCountSeason <- 
  filter( cd, !is.na( riverOrderedLagged ) ) %>%
  count(riverOrdered,species,season) 

rrCountSeason <- 
filter( cd, !is.na( riverOrderedLagged ) ) %>%
  group_by(riverOrdered, riverOrderedLagged, species, season) %>%
  tally() %>%
  left_join( .,rCountSeason, by = c("riverOrdered","species","season") ) %>%
  mutate( prop = n.x / n.y ) %>%
  complete( riverOrdered,riverOrderedLagged,species,season ) %>%  # not sure why this adds 4* the # of rows
  distinct() 
  #mutate( n.y = ifelse( is.na( n.y ), 0, n.y ) )
  
propMoveSeason <- 
  ggplot( rrCountSeason, aes( x=riverOrderedLagged, y=prop ) ) +
    geom_bar( stat = "identity" ) +
    scale_fill_grey(name = "Season") +
    scale_x_discrete('Second river') +
    scale_y_continuous("Proportion") +
    theme_bw(base_size=25) +
    theme( legend.position=c(0.95,0.9) ) +
    geom_text( aes(label=(n.y), x=4.2,y=1.025 ), size=6, family="sans" ) +
    facet_grid( species ~ riverOrdered + season )  

  ggsave( file=paste0(graphsDir,'propMoveSeason.png'), plot=propMoveSeason, dpi=600 , width=9,height=5, units='in', scale=2 )
  

```

```{r counts and biomasses by section}

# Nested residuals (for means) and proportions (for counts)  

bySn <-   cd %>%
  group_by(sampleNumber) %>%
  summarize( bioBySnSum = sum( observedWeight, na.rm = T ),
             bioBySnMean = mean( observedWeight, na.rm = T ),
             countBySn = n() )
  
bySnRivSec <- cd  %>%
       group_by( sampleNumber, riverOrdered, section ) %>%
       summarize( bioBySnRivSecSum = sum( observedWeight, na.rm = T  ),
                  bioBySnRivSecMean = mean( observedWeight, na.rm = T  ),
                  countBySnRivSec = n() ) %>%
       left_join( .,bySn, by = c("sampleNumber") ) %>%
       mutate( bioBySnRivSecResid = bioBySnRivSecMean - bioBySnMean,
               countBySnRivSecProp = countBySnRivSec / countBySn )

bySnRivSecSpp <- cd  %>%
       group_by( species, sampleNumber, riverOrdered, section ) %>%
       summarize( bioBySnRivSecSppSum = sum( observedWeight, na.rm = T  ),
                  bioBySnRivSecSppMean = mean( observedWeight, na.rm = T  ),
                  countBySnRivSecSpp = n() ) %>%
       left_join( .,bySnRivSec, by = c("sampleNumber", "riverOrdered", "section") ) %>%
       mutate(   bioBySnRivSecSppResid =  bioBySnRivSecSppMean - bioBySnRivSecMean,
               countBySnRivSecSppProp = countBySnRivSecSpp /   countBySnRivSec )

#Check data  
#ggplot( bySnRivSecSpp, aes( countBySnRivSecSpp,bioBySnRivSecSppSum, color=section) ) + geom_point() + facet_grid(riverOrdered~species)  
 
cd2 <- left_join( cd, bySnRivSecSpp, by = c("species","sampleNumber", "riverOrdered", "section") ) %>%
  mutate( moved = !( section == sectionLagged & riverOrdered == riverOrderedLagged ) )
         

```

```{r biomass and movement}
#ggplot( cd2, aes( bioBySnRivSecSppResid, moveDist ) ) +
ggplot( cd2, aes( bioBySnRivSecResid, moveDist ) ) +
  geom_point() +
  geom_smooth() +
  facet_grid( riverOrdered~species )

ggplot( cd2, aes( moved,bioBySnRivSecSppResid ) ) +
  geom_boxplot() +
  facet_grid( species ~ riverOrdered )

#m0 <- glm( moved ~ biomassBySection*species*riverOrdered, family="binomial",data=cd2)

```

```{r body size and movement}

ggplot( cd2, aes( moved,observedLength ) ) +
  geom_boxplot() +
  facet_grid( species ~ riverOrdered )

ggplot( filter(cd2,!is.na(season) & observedLength < 200  ), aes( observedLength, color=moved ) ) +
  geom_freqpoly() +
  facet_grid( species + season ~ riverOrdered, scales = 'free_y' )

binomial_smooth <- function(...) {
   geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
 }

ggplot( filter(cd2,!is.na(season) & observedLength < 200  ), aes( x=observedLength, y=as.numeric(moved ), color=species ) ) +
  geom_point(alpha = 0.05) +
  binomial_smooth() +
  facet_wrap( ~ riverOrdered )
ggplot( filter(cd2,!is.na(season) & observedLength < 200  ), aes( x=observedLength, y=as.numeric(moved ), color=riverOrdered ) ) +
  geom_point(alpha = 0.05) +
  binomial_smooth() +
  facet_wrap( ~ species )

ggplot( filter(cd2,!is.na(season) & observedLength < 200 & moved ), aes( x=observedLength, y=moveDist, color=riverOrdered ) ) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap( riverOrdered~ species )

ggplot( filter(cd2,!is.na(season) & observedLength < 200  ), aes( x=observedLength, y=as.numeric(moved ) ) ) +
  geom_point(alpha = 0.05) +
  binomial_smooth() +
  facet_grid( season ~ species + riverOrdered )

length, density-leave, density-stop, season, river, species
```

